### 2016-11-30 thesis/intro/intro.tex
tmp commit

With that said, there has been a few environmental factors that have a strong evidence of causing cancer, though the complete mechanism of action for some of these must be elucidated in the coming years.
% TODO: do I really need infections and inflammations?
Out of these, infections and inflammation, smoking, \gls{uv} radiation, and diet and obesity will be covered as examples.
\\

% Infections and inflammmation

@@

\noindent
As one can imagine, \acrshort{dna} damages caused by smoking and/or \gls{uv} radiation have the potential to damage the genome and initiate cancer if wrong genes are mutated, such as p53 tumour suppressor gene.
Therefore, limited exposure to both factors is crucial for the prevention of any types of cancers, as well as the more specific types of cancers associated with these factors, such as lung and skin cancers.\\

% Diet and obesity
\noindent

>>>>

With that said, there has been a few environmental factors that have a strong evidence of causing cancer, though the complete mechanism of action for some of these must be elucidated in the coming years.
Out of these, smoking, \gls{uv} radiation, and diet and obesity will be covered as examples.
\\

@@

% Diet and obesity
\noindent
Finally, diet and obesity are known risk factors for many types of cancers \citep{Ames1995a,Calle2004}.
Though more will be covered in details in \cref{sec:obesity_and_cancers}: \nameref{sec:obesity_and_cancers}, the effect of obesity on cancers will be briefly mentioned here.

Epidemiological evidence of obesity being a risk factor for many cancers has been well established, yet the mechanisms of its contribution to cancer is still  speculative \citep{Calle2003,Kelesidis2006a}.
Although there are a lot of possible mechcanisms, there is growing evidence of the role of chronic inflammation caused by increased fat mass in the body \citep{Kelesidis2006a,Lumeng2011,Hernandez2013}.
Free fatty acids from excess adipose tissue and/or diet can be recognised by the \gls{tlr}4, which can then activate \gls{nfkb} \citep{Lumeng2011}.
Clearly, the result of the activation of \gls{nfkb} can lead to the activation of one of the Hallmarks of Cancers -- tumour-promoting inflammation.

The nutritional content of the food that people consume is also important, as it can both directly and indirectly influence the likelihood of tumorigenesis, as well as its progression \citep{Ames1995a}.
Many fruits and vegetables contain antioxidants as well as essential viatmins and minerals.
Antioxidants are able to neutralise the effect of \gls{ros},  and many viatmins have been associated with reduction of many types of cancers \citep{Ames1995a}.
In fact, ascorbic acid (vitamin C) is able to prevent the lipid peroxidation of lipid plasma caused by tobacco smokes \textit{in vitro}, which could be why smokers, compared to non-smokers, have decreased levels of ascorbic acid \citep{Hecht1999}.

By maintaining healthy diet, it may be possible to reduce the risks of many cancers and obesity, as well as the comorbidities associated with them.
\\

\noindent
As one can imagine, \acrshort{dna} damages caused by smoking and/or \gls{uv} radiation have the potential to damage the genome and initiate cancer if wrong genes are mutated, such as p53 tumour suppressor gene.
Healthy diet is also crucial for lowering the risk of cancers through beneficial vitamins and antioxidants, and prevent obesity as well as chronic inflammation that accompanies with it.
There are many other environmental factors associated with cancers, and limiting the exposure to detrimental environmental factors is crucial for the prevention of any types of cancers, as well as the more specific types of cancers associated with these factors, such as lung and skin cancers.
\\
